---
title: "Average Home Purchasing Price Analysis"
author: "Lianghui Li, River He, Zhongshu Wang, Boyue Sun"
date: "December 6, 2015"
output: pdf_document
---

```{r, echo=F, warning=F, include=F}
source('../code/Lianghui.r')
source('../code/River&Boyue.r')
source('../code/Zhongshu.r')
```

```{r, include=F}
#this part only be included because of problems of export PDF files from ggplot  in original code files

pdf("../images/sf_rate2001_graph.pdf")
sf_rate2001_graph
dev.off()

pdf("../images/housing_index.pdf")
housing_index_graph
dev.off()

png("../images/housing_index.png")
housing_index_graph
dev.off()

pdf("../images/time_trend.pdf")
time_trend
dev.off()
```


#DATA ANALYSIS

##Housing Affordbility(Lianghui Li)

Colletcting interest rate(per month)  from"data360.org"   
            income distribution(per year) from "census.gov"   
            and housing price(per month) from "zillow"


###Time Trend

compare the price trend of minimum average price _WV_ and maximum average price _HI_

```{r,echo=F, tidy=T}
time_trend
```

------

###General Question

__How much income to apply for mortgage?__

qualify income to apply mortgage in the year 2003, 2013 in state _AK_ & _CA_
```{r, warning=F}
qualify_income('AK', 2003)
qualify_income('CA', 2013)
```

__Your income is high enough?__

housing index for the income value 64930 in year 2008, 2010 in state _AL_ & _IL_
```{r, warning=F}
housing_index(64930, 'AL', 2009)
housing_index(64930, 'DC', 2010)
#notice that if the year is not between 1996 and 2013, no result will be shown
qualify_income('NJ', 1993)
```

__Is majority of American can afford?__

overall American affordbility for the housing price in _OK_ & _CT_ in 1999 and 2001
```{r, warning=F}
housing_affordability('OK', 1999)
housing_affordability('CT', 2008)
```

-----

##General Affordbility


Bar plot that show affordbility for each income distribution in the year 2008 for each state

```{r, echo=F,tidy=T}
housing_index_graph
```

From the graph, we can see that expensive housing states for general American are HI,DC, CA; and cheapest are WV, OK and NE

_Limitation_ : Since our formula just include two elements(interest rate and   income); however, general concept of affordabilty is not limited, but also   contains factor such as __number of school__, __transportation cost__,   
__local tax policy__, etc. These omitted variable could bias our calculating the affordability index.  


------

#Realtionship between Housing Price and Education(River He & Boyue) 
_collecting data from "ftp.cde.ca.gov" and "web.alsde.edu" 
However, when we first downloaded the data from "web.alsde.edu",we could use
it, but somehow the website link has some problems when we use downloadfile     command.So we just put AlabamaSchoolListing.csv in our raw data folder_
  
## relationship between housing price and number of schools in Alabama's counties


```{r,echo=F}
#use the prices in oct 2015 to plot (linear model)
plot(New_merged_Al[,242]~New_merged_Al[,243],main = "Price of 2015.10 VS Number of Schools",
     xlab = "Schools",ylab ="Price",pch=11, cex=1.1, col="#33BB79")
```

```{r,echo=F}
cor(New_merged_Al[,242],New_merged_Al[,243])
```
There is a positive relationship between house price and school numbers in Alabama but the correlation is only 0.019 which is very small. We can ignore it. 
Also from this graph, we found the relationship is not linear,so we decided to use linear-log model.

```{r,echo=F}
# graphs use linear-log model
plot(New_merged_Al[,242]~log(New_merged_Al[,243]),
     main = "Price of 2015.10 VS log form of School numbers",
     xlab = "log of School #",ylab ="Price",pch=13, cex=1.2, col="#2929DD")
```

```{r,echo=F}
cor(New_merged_Al[,242],log(New_merged_Al[,243]))
```
There is a positive relationship between house price and school numbers in Alabama and the correlation is only 0.3374 which is higher. 

## relationship between housing price and number of schools in California's counties
```{r, echo=F}
# use the prices in oct 2015 to plot (linear model)
plot(
  Ca[,242] ~ Ca[,243],
  main = "Price of 2015.10 VS Number of Schools",
  xlab = "Schools",ylab = "Price",cex = 1,pch = 7,col = hsv(h = 0.9,s =
                                                              0.7,v = 0.8)
)

```

```{r}
cor(Ca[,242],Ca[,243])
```
There is a positive relationship between housing price and number of schools in California's counties. However, the correlation is only 0.19, it is really small. 



```{r, echo=F}
#linear-log model
plot(
  Ca[,242] ~ log(Ca[,243]),
  main = "Price of 2015.10 VS log form of School numbers",
  xlab = "log of School #",ylab = "Price",
  cex = 1,pch = 9 ,col = hsv(h = 0.5,s = 0.5,v = 0.5)
)
```

```{r}
cor(Ca[,242],log(Ca[,243]))
```
After doing the linear model, we found that the relationship between housing price and number of schools is not linear, so we decided to use linear-log model to construct the graph. 
There is a positive relationship between housing price and number of schools in California's counties. The correlation in linear-log model is 0.36 which is larger than the correlation we got before. 


_Limitation_ : It is hard to measure the qualities of schools, which may have impact on the housing price. Also we only choose two states in this part, the results may be different from other states.
-----------

## Housing Price and Stock Market(Zhongshu Wang)
collect average housing price from zillow research database
        s&p500 index from yahoo finance with preselecting month information
Take San Francisco as an example, comparing housing price and s&p500.
Here, we divided the average housing price by 500 to be more compariable


```{r,echo=F, warning=F}
plot(
    1:nrow(sf_index),sf_index$house / 500, col = "#EE0000", type = "l", lwd = 3,at = FALSE,xlab = "Months",ylab = "Price Index" 
  )
  lines(
    1:nrow(sf_index),sf_index$sp, col = "#00EE00", lwd = 3,at = FALSE
  )
  axis(side = 1,at = seq(10,nrow(sf_index),by = 12),labels = FALSE)
  axis(side=2, at=seq(500,2300,by=200),labels=seq(500,2300,by=200),las=1)
  text(seq(10,nrow(sf_index),by = 12),par("usr")[3] - 0.06,cex = 0.8, 
       labels = sf_index$time[seq(10,nrow(sf_index),by =12)],
       srt= 45, pos = 1, xpd = TRUE)
  legend("topleft",legend=c("Housing Price / 500","S&P 500"),col = c("#EE0000","#00EE00"),lty = c(1,1))
  title("Price Comparison")
```

However, this graph can't be develeped into a strong argument

Real Estate and Stock Markets are two investment viechles, which they are subsitutes.
Also, Real Estate might affect the stock market as a factor.

Thus, the Dot Com bubble bust in 2001 caused the stock market downtrend, investors lost confidents and looking for other investment viechles.
Hence, we conduct an observation of the relationship between housing purchase price and stock price on their rate of change.

Here is the graph of their rate of change during 2001

```{r,echo=F, tidy=T, message=FALSE, comment=FALSE}
sf_rate2001_graph

```
Graphs shows that rate of change on housing price moved in the opposite direction as the rate of change on s&p 500
Moreover, here is the correlation test
```{r}
cor(sf_rate2001$house_rate,sf_rate2001$sp_rate)
```
There is a week negative correlation between housing price and stock price in 2001
```{r}
cor(sf_rate$house_rate,sf_rate$sp_rate)
```
But, the longrun was more likely to be uncorrelated because other factors which influnced behaviors in both markets 

_Limitation_ : There are too many confounding variables that influence both market either into the same direction or the different direction. Our research is only based on observation on short term period and a specific region, we don't have sufficient information and skills to do the full economical analysis.